### 2013-07-01 floats/tables.tex
Physically moved floats to the end

% floats/tables.tex

% tab:reanalyses
\begin{table*}[t]
	\caption{Characteristic of reanalysis climatologies used to force the ice sheet model.}
	\label{tab:reanalyses}
	\vskip4mm
	\centering
	\begin{tabular}{lllll}
		\tophline
		Reanalysis& Spatial coverage& Averaging period& Resolution& Description\\
		\middlehline
		NCEP/NCAR&  global&     1981 -- 2010& 1.875\degree& \citet{data:ncar}\\
		ERA-Interim&global&     1979 -- 2011& 1.000\degree& \citet{data:erai}\\
		CFSR&       global&     1979 -- 2010& 0.325\degree& \citet{data:cfsr}\\
		NARR&       North America& 1979 -- 2000& 32\,km& \citet{data:narr}\\
		\bottomhline
	\end{tabular}
\end{table*}