### 2013-07-01 sections/concl.tex

% section/concl.tex
% ----------------------------------------------------------------------
\conclusions
\label{sec:concl}
% ----------------------------------------------------------------------

\emph{Coming soon...}