### 2013-09-14 sections/abstract.tex
Implement corrections from Qiong

% section/abstract.tex
% ----------------------------------------------------------------------

\begin{abstract}
We present an ensemble of numerical simulations of the Cordilleran Ice Sheet during the Last Glacial Maximum performed with the Parallel Ice Sheet Model (PISM), using temperature offsets from present-day climatologies from five different datasets. Monthly mean surface air temperature and precipitation from the NCEP/NCAR reanalysis, the Climate Forecast System Reanalysis, the ERA-Interim reanalysis and the North American Regional Reanalysis are used to compute surface mass balance by using a positive degree-day model. Modelled ice sheet outlines and volumes appear highly sensitive to the choice of climate forcing. We assess model performance against a reconstruction of the ice margin at the Last Glacial Maximum in continental regions of northern Yukon Territory and interior Alaska. The best match between model output and the reconstructed ice margin is obtained using climate forcing from the North American Regional Reanalysis, which is a regional climate reanalysis with the highest resolution.
\end{abstract}