### 2013-11-05 floats/figures.tex
Add new figure captions

% floats/figures.tex
% ----------------------------------------------------------------------

% fig:locmap
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics[width=8cm]{cordillera-climate-locmap}
	\end{center}
	\caption{Shaded relief map of northern North America with the extent of the Cordilleran (CIS), Laurentide (LIS) and Innuitian (IIS) ice sheets at 14\,$^{14}$C\,ka\,BP (16.8\,cal\,ka\,BP) \citep{dyke-2004}. The background map consists of ETOPO1 \citep{data:etopo1} and Natural Earth Data \citep{data:naturalearth}. Assembled with GRASS~GIS \citep{soft:grass} and Inkscape \citep{web:inkscape}}
	\label{fig:locmap}
\end{figure}

% fig:topo
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-topo}
	\end{center}
	\caption{Topography maps from WorldClim, ERA-Interim reanalysis, NARR (North American Regional Reanalysis), CFSR (Climate System Forecast Reanalysis), and NCEP/NCAR reanalysis climatologies used as a reference for temperature lapse-rate corrections, and topography map from ETOPO1 used as basal topography. Fig.~\ref{fig:topo}-\ref{fig:best} drawn using Matplotlib \citep{soft:mpl}.}
	\label{fig:topo}
\end{figure*}

% fig:temp
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-temp}
	\end{center}
	\caption{Summer (JJA) and winter (DJF) temperature maps from WorldClim, and summer (JJA) temperature maps from ERA-Interim reanalysis, NARR (North American Regional Reanalysis), CFSR (Climate System Forecast Reanalysis), and NCEP/NCAR reanalysis climatologies.}
	\label{fig:temp}
\end{figure*}

% fig:prec
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-prec}
	\end{center}
	\caption{Winter (DJF) and summer (JJA) precipitation maps from WorldClim, and winter (DFJ) temperature maps from ERA-Interim reanalysis, NARR (North American Regional Reanalysis), CFSR (Climate System Forecast Reanalysis), and NCEP/NCAR reanalysis climatologies. 	Additional forcing data was prepared to correct for wave-like precipitation artefacts in CFSR (section~\ref{sec:climate}).}
	\label{fig:prec}
\end{figure*}

% fig:ivolarea
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-ivolarea}
	\end{center}
	\caption{Total glaciated area and ice volume after 10\,kyr as a function of temperature offset and climate forcing.}
	\label{fig:ivolarea}
\end{figure}

% fig:extent
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-extent}
	\end{center}
	\caption{Extent of ice cover after 10\,kyr as a function of applied temperature offsets for each climate forcing.}
	\label{fig:extent}
\end{figure*}

% fig:cool05
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-cool05}
	\end{center}
	\caption{Ice surface topography (black contours every 1000\,m) and velocity (\unit{m\,yr^{-1}}) after 10\,kyr under a climate 5\,\unit{\degree C} colder than present for each climate forcing.}
	\label{fig:cool05}
\end{figure*}

% fig:tempheatmap
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-tempheatmap}
	\end{center}
	\caption{Logarithmically coloured density maps, showing a comparison of JJA surface air temperature data from the WorldClim climatology, against that of each atmospheric reanalysis. Apparent horizontal lines are an artefact of different horizontal resolutions. Note the cold bias of NCEP/NCAR data.}
	\label{fig:tempheatmap}
\end{figure}

% fig:tempdiff
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-tempdiff}
	\end{center}
	\caption{JJA surface air temperature difference maps against WorldClim data, after bi-linear spatial interpolation. Note the cold bias of NCEP/NCAR data and temperature anomalies due to unresolved topographic detail.}
	\label{fig:tempdiff}
\end{figure}

% fig:precheatmap
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-precheatmap}
	\end{center}
	\caption{Logarithmically coloured density maps, showing a comparison of DJF precipitation rate from the WorldClim climatology, against that of each atmospheric reanalysis. Apparent horizontal lines are an artefact of different horizontal resolutions. Apparent vertical lines at low values are an artefact of WorldClim data precision. Note the wet bias of all reanalysis data.}
	\label{fig:precheatmap}
\end{figure}

% fig:precdiff
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-precdiff}
	\end{center}
	\caption{DJF precipitation rate relative difference maps against WorldClim data, after bi-linear spatial interpolation. Note the wet bias of all reanalysis data and the large anomalies of CFSR and NCEP/NCAR data.}
	\label{fig:precdiff}
\end{figure}

% fig:oroprecip
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-oroprecip}
	\end{center}
	\caption{Schematic representation of the orographic precipitation effect over the Coast Mountains. In a GCM of lower resolution, the precipitation peak appears smoother and the precipitation shadow is not as pronounced. Drawn using Inkscape \citep{web:inkscape}.}
	\label{fig:oroprecip}
\end{figure}

% fig:biatm
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-biatm}
	\end{center}
	\caption{“Hybrid” atmosphere simulations using precipitation rate from WorldClim and surface air temperature from each reanalysis (upper row), and vice-versa (lower row). In other words, the upper row shows the effect of temperature anomalies, and the lower row the effect of precipitation anomalies, for each reanalysis, relative to WorldClim data. Each simulation use a 5\,\unit{\degree C} offset. Compare with figure~\ref{fig:cool05}}
	\label{fig:biatm}
\end{figure*}

% fig:biatmbars
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-biatmbars}
	\end{center}
	\caption{Effect of temperature and precipitation anomalies of each reanalysis, and both, on modelled final ice volume, using the WorldClim 5\,\unit{\degree C} offset simulation as a reference. Corresponding ice sheet geometries are presented in figures~\ref{fig:cool05} and~\ref{fig:biatm}}
	\label{fig:biatmbars}
\end{figure*}

% fig:best
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-best}
	\end{center}
	\caption{Ice surface topography (black contours every 1000\,m) after 10\,kyr using temperature offsets that lead to similar areas of ice cover for each climate forcing. 14\,$^{14}$C\,ka\,BP (16.8\,cal\,ka\,BP) ice margin (blue line) from \citet{dyke-2004}.}
	\label{fig:best}
\end{figure*}

% fig:durationstack
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-durationstack}
	\end{center}
	\caption{Model sensitivity to simulation length. Modelled glaciated area, using NARR forcing data, and temperature offsets from 0 to 15\,\unit{\degree C}, solid lines corresponding to values from 7 to 11\,\unit{\degree C} (left). Corresponding modelled ice margin, compared to the LGM ice sheet margin by \citep{dyke-2004} (right).}
	\label{fig:durationstack}
\end{figure}