### 2013-11-06 cordillera-climate.tex
Add authors

% cordillera-climate.tex
% ----------------------------------------------------------------------

% Copernicus manuscript
\documentclass[tc, ms]{copernicus}

% Copernicus final print
%\documentclass[tc]{copernicus}

% Copernicus discussion paper
%\documentclass[tcd, hvmath]{copernicus_discussions}

% Copernicus-like latex2rtf compatible
%\input{copernicus_rtf}

% Coloured hyperlinks
\usepackage[colorlinks,citecolor=blue]{hyperref}

% Recognize utf-8 special characters
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

% To-do footnotes
\newcommand{\needref}[0]{\footnote{Reference needed}}
\newcommand{\julien}[1]{\footnote{\textbf{Julien:} #1}}
\newcommand{\arjen}[1]{\footnote{\textbf{Arjen:} #1}}
\newcommand{\qiong}[1]{\footnote{\textbf{Qiong:} #1}}
\newcommand{\irina}[1]{\footnote{\textbf{Irina:} #1}}

% Set figures repository
\graphicspath{{figures/}}

% ----------------------------------------------------------------------
\begin{document}
% ----------------------------------------------------------------------

% Title
\title{The effect of climate forcing on numerical simulations of the Cordilleran Ice Sheet at the Last Glacial Maximum}
\runningtitle{Climate forcing for Cordilleran ice sheet simulations}

% Authors
\author[1]{Julien Seguinot}
\author[2]{Constantine Khroulev}
\author[3]{Irina Rogozhina}
\author[1]{Arjen Stroeven}
\author[1]{Qiong Zhang}
\runningauthor{J. Seguinot et al.}
\correspondence{J. Seguinot\\ (julien.seguinot@natgeo.su.se)}

% Affiliations
\affil[1]{Department of Physical Geography and Quaternary Geology, Stockholm University, Stockholm, Sweden}
\affil[2]{Geophysical Institute, University of Alaska Fairbanks, Fairbanks, AK, USA}
\affil[3]{Helmholtz Centre Potsdam, GFZ German Research Centre for Geosciences, Potsdam, Germany}

% For Copernicus
\received{}
\pubdiscuss{}
\revised{}
\accepted{}
\published{}

% Title
\maketitle

% Sections
\input{sections/abstract}
\input{sections/intro}
\input{sections/model}
\input{sections/climate}
\input{sections/results}
\input{sections/discuss}
\input{sections/concl}

% Acknowledgements
%\begin{acknowledgements}
%\end{acknowledgements}

% References
\bibliographystyle{copernicus}
\bibliography{references}

% Floats
\newpage
\input{floats/tables}
\input{floats/figures}

% ----------------------------------------------------------------------
\end{document}
% ----------------------------------------------------------------------