### 2013-11-07 floats/figures.tex
Arjen corrections in figure captions

% floats/figures.tex
% ----------------------------------------------------------------------

% fig:locmap
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics[width=8cm]{cordillera-climate-locmap}
	\end{center}
	\caption{Shaded relief map of northern North America with the extent of the Cordilleran (CIS), Laurentide (LIS) and Innuitian (IIS) ice sheets at 14\,$^{14}$C\,ka\,BP (16.8\,cal\,ka\,BP) \citep{dyke-2004}. While this age denotes the LIS after retreat from its LGM, it closely corresponds to the LGM extent of most of the Cordilleran ice sheet \citep{porter-swanson-1998,dyke-2004,stroeven-etal-2010}. The rectangular box denotes the modelling domain of the Cordilleran ice sheet of 1500 by 3000\,km. The background map consists of ETOPO1 \citep{data:etopo1} and Natural Earth Data \citep{data:naturalearth} and was assembled with GRASS~GIS \citep{soft:grass}.}
	\label{fig:locmap}
\end{figure}

% fig:topo
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-topo}
	\end{center}
	\caption{Topography maps from WorldClim \citep{data:worldclim}, ERA-Interim reanalysis \citep{data:erai}, North American Regional Reanalysis (NARR) \citep{data:narr}, ETOPO1 \citep{data:etopo1}, Climate System Forecast Reanalysis (CFSR) \citep{data:cfsr}, and NCEP/NCAR reanalysis \citep{data:ncar}. Whereas ETOPO1 is used as basal topography for the ice sheet model, all others are used data used as a reference for temperature lapse-rate corrections. Fig.~\ref{fig:topo}-\ref{fig:best} drawn using Matplotlib \citep{soft:mpl}.}
	\label{fig:topo}
\end{figure*}

% fig:temp
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-temp}
	\end{center}
	\caption{Summer (JJA) and winter (DJF) temperature maps from WorldClim \citep{data:worldclim}, and summer (JJA) temperature maps from ERA-Interim reanalysis \citep{data:erai}, North American Regional Reanalysis (NARR) \citep{data:narr}, Climate System Forecast Reanalysis (CFSR) \citep{data:cfsr}, and NCEP/NCAR reanalysis \citep{data:ncar} climatologies.}
	\label{fig:temp}
\end{figure*}

% fig:prec
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-prec}
	\end{center}
	\caption{Winter (DJF) and summer (JJA) precipitation maps from WorldClim \citep{data:worldclim}, and winter (DFJ) temperature maps from ERA-Interim reanalysis \citep{data:erai}, North American Regional Reanalysis (NARR) \citep{data:narr}, Climate System Forecast Reanalysis (CFSR) \citep{data:cfsr}, and NCEP/NCAR reanalysis \citep{data:ncar} climatologies. Additional forcing data was prepared to correct for wave-like precipitation artefacts in CFSR (section~\ref{sec:climate}).}
	\label{fig:prec}
\end{figure*}

% fig:ivolarea
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-ivolarea}
	\end{center}
	\caption{Total glaciated area and ice volume after 10\,kyr as a function of temperature offset and climate forcing used.}
	\label{fig:ivolarea}
\end{figure}

% fig:extent
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-extent}
	\end{center}
	\caption{Extent of ice cover after 10\,kyr as a function of applied temperature offsets for each climate forcing used.}
	\label{fig:extent}
\end{figure*}

% fig:cool05
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-cool05}
	\end{center}
	\caption{Ice surface topography (1 km contours) and velocity (\unit{m\,yr^{-1}}) after 10\,kyr under a climate 5\,\unit{\degree C} colder than present for each climate forcing used.}
	\label{fig:cool05}
\end{figure*}

% fig:tempheatmap
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-tempheatmap}
	\end{center}
	\caption{Logarithmically coloured density maps, showing a comparison of JJA surface air temperature data from the WorldClim climatology, against that of each atmospheric reanalysis. Apparent horizontal lines are an artefact of different horizontal resolutions. Note the cold bias of NCEP/NCAR data.}
	\label{fig:tempheatmap}
\end{figure}

% fig:tempdiff
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-tempdiff}
	\end{center}
	\caption{JJA surface air temperature difference maps against WorldClim data, after bi-linear spatial interpolation. Note the cold bias of NCEP/NCAR data and temperature anomalies due to unresolved topographic detail.}
	\label{fig:tempdiff}
\end{figure}

% fig:precheatmap
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-precheatmap}
	\end{center}
	\caption{Logarithmically coloured density maps, showing a comparison of DJF precipitation rate from the WorldClim climatology, against that of each atmospheric reanalysis. Apparent horizontal lines are an artefact of different horizontal resolutions. Apparent vertical lines at low values are an artefact of WorldClim data precision. Note the wet bias of all reanalysis data.}
	\label{fig:precheatmap}
\end{figure}

% fig:precdiff
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-precdiff}
	\end{center}
	\caption{DJF precipitation rate relative difference maps against WorldClim data, after bi-linear spatial interpolation. Note the wet bias of all reanalysis data and the large anomalies of CFSR and NCEP/NCAR data.}
	\label{fig:precdiff}
\end{figure}

% fig:oroprecip
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-oroprecip}
	\end{center}
	\caption{Schematic representation of the orographic precipitation effect over a mountain range. In a GCM of low resolution, the precipitation peak appears downwind-shifted and smoother and the precipitation shadow is less pronounced than in a high resolution GCM.}
	\label{fig:oroprecip}
\end{figure}

% fig:biatm
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-biatm}
	\end{center}
	\caption{“Hybrid” atmosphere simulations using precipitation rate from WorldClim and surface air temperature from each reanalysis (upper row), and vice-versa (lower row).  (1 km contours)In other words, the upper row shows the effect of temperature anomalies, and the lower row the effect of precipitation anomalies, for each reanalysis, relative to WorldClim data. Each simulation use a 5\,\unit{\degree C} offset. Compare with figure~\ref{fig:cool05}}
	\label{fig:biatm}
\end{figure*}

% fig:biatmbars
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-biatmbars}
	\end{center}
	\caption{Effect of temperature and precipitation anomalies of each reanalysis, and both, on modelled final ice volume, using the WorldClim 5\,\unit{\degree C} offset simulation as a reference. Corresponding ice sheet geometries are presented in figures~\ref{fig:cool05} and~\ref{fig:biatm}}
	\label{fig:biatmbars}
\end{figure}

% fig:best
\begin{figure*}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-best}
	\end{center}
	\caption{Ice surface topography (1\,km contours) after 10\,kyr using temperature offsets resulting in glaciated areas of circa $2\,\times10^6\,\unit{km^2}$. A reconstructed LGM ice sheet margin by \citet{dyke-2004} (Fig.~\ref{fig:locmap}) is given for reference (blue line).}
	\label{fig:best}
\end{figure*}

% fig:durationstack
\begin{figure}[t]
	\vspace*{2mm}
	\begin{center}
		\includegraphics{cordillera-climate-durationstack}
	\end{center}
	\caption{Model sensitivity to simulation length. Modelled glaciated area, using NARR forcing data, and temperature offsets from 0 to 15\,\unit{\degree C}, solid lines corresponding to values from 7 to 11\,\unit{\degree C} (left). Corresponding modelled ice margin, compared to the LGM ice sheet margin by \citep{dyke-2004} (right).}
	\label{fig:durationstack}
\end{figure}