### 2013-11-19 cordillera-climate.tex
Add acknowledgements

% cordillera-climate.tex
% ----------------------------------------------------------------------

% Copernicus manuscript
\documentclass[tc, ms]{copernicus}

% Copernicus final print
%\documentclass[tc]{copernicus}

% Copernicus discussion paper
%\documentclass[tcd, hvmath]{copernicus_discussions}

% Copernicus-like latex2rtf compatible
%\input{copernicus_rtf}

% Coloured hyperlinks
\usepackage[colorlinks,citecolor=blue]{hyperref}

% Recognize utf-8 special characters
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

% Set figures repository
\graphicspath{{figures/}}

% ----------------------------------------------------------------------
\begin{document}
% ----------------------------------------------------------------------

% Title
\title{The effect of climate forcing on numerical simulations of the Cordilleran Ice Sheet at the Last Glacial Maximum}
\runningtitle{Climate forcing for Cordilleran ice sheet simulations}

% Authors
\author[1]{Julien Seguinot}
\author[2]{Constantine Khroulev}
\author[3]{Irina Rogozhina}
\author[1]{Arjen P. Stroeven}
\author[1]{Qiong Zhang}
\runningauthor{J. Seguinot et al.}
\correspondence{J. Seguinot\\ (julien.seguinot@natgeo.su.se)}

% Affiliations
\affil[1]{Department of Physical Geography and Quaternary Geology, Stockholm University, Stockholm, Sweden}
\affil[2]{Geophysical Institute, University of Alaska Fairbanks, Fairbanks, AK, USA}
\affil[3]{Helmholtz Centre Potsdam, GFZ German Research Centre for Geosciences, Potsdam, Germany}

% For Copernicus
\received{}
\pubdiscuss{}
\revised{}
\accepted{}
\published{}

% Title
\maketitle

% Sections
\input{sections/abstract}
\input{sections/intro}
\input{sections/model}
\input{sections/climate}
\input{sections/results}
\input{sections/discuss}
\input{sections/concl}

% Acknowledgements
\begin{acknowledgements}
We thank Ed Bueler and Andy Aschwanden for providing constant support on PISM and patiently answering our questions on the code internals, Laurent Brodeau and Soon-Heum Ko for assistance concerning software installations on supercomputing resources, and all PISM authors for providing the model open-source. Funding was provided by the Swedish Research Council (VR) to A. Stroeven (No. 2008-3449) and Stockholm University. Computer resources were provided by the Swedish National Infrastructure for Computing (SNIC) at the National Supercomputing Center (NSC) to A. Stroeven (No. 2013/1-145).

% Author contributions
\textit{Author contributions.}
J. Seguinot ran the simulations and wrote most of the paper; C. Khroulev implemented necessary code changes in PISM; I. Rogozhina helped with model set-up and experiment design; A. Stroeven initiated the Cordilleran Ice Sheet modelling project; and Q. Zhang prepared some of the forcing data. All authors contributed to the redaction.
\end{acknowledgements}

% References
\bibliographystyle{copernicus}
\bibliography{references}

% Floats
\newpage
\input{floats/tables}
\input{floats/figures}

% ----------------------------------------------------------------------
\end{document}
% ----------------------------------------------------------------------