### 2013-12-05 reviews/initial-response.tex
Address minor editor comments

\documentclass{letter}

\signature{Julien Seguinot}
\address{
  julien.seguinot@natgeo.su.se\\\\
  Stockholm University\\
  SE-106 91 Stockholm\\
  SWEDEN}

\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{xcolor}
%\usepackage{enumitem}
%\usepackage{letterbib}

\newcommand{\rev}[0]{\color{blue!50!black}\it}
\newcommand{\textrev}[1]{{\rev``#1''}}
\newcommand{\revpoint}[1]{{\rev\item``#1''}}
\newcommand{\todo}[1]{\textcolor{red!50!black}{Todo: #1}}
\newcommand{\done}[1]{\textcolor{green!50!black}{Done: #1}}

% ----------------------------------------------------------------------

\begin{document}
\begin{letter}{Reply to Editor Initial Decision}

\opening{Dear Editor,}

Many thanks for accepting our paper for publication in TCD. We address your comments in the updated manuscript and detail the changes made below.

% ----------------------------------------------------------------------

\begin{itemize}

\revpoint{Line 43: please rephrase this sentence: 'other side of the coin' and 'hand-in-hand' into something of higher scientific value.}

This sentence was rephrased to: 'However, avoiding this circular dependence requires simplifying assumptions regarding Earth's past climate.'

\revpoint{Line 47: idem: Standing on middle ground is not a good beginning of the sentence. What is actually meant by this. Be more precise.}

Rephrased to 'Alternatively.'

\revpoint{Line 59-60: remove 'In our study'. just start sentence by: 'To limit the degrees of freedom ...'}

Rephrased accordingly.

\revpoint{Line 85: remove 'Thereby'. Just start with 'We aim to determine ...'}

Rephrased accordingly.

\revpoint{Line 87-88: Remove this sentence. It may be the first study to focus on the Cordilleran ice sheet since Robert (1991), but isn't that at the same time a weakness, since the model is not taking into account the buttressing effect of the Laurentide, which is done in other model studies that take into account the whole North American continent?}

This sentence was removed.

\revpoint{Line 113-116: Is this realistic to take a constant value of 70 $mW/m^2$? Aren't there direct measurements in the region to support this. Since it is an active margin, there could well be quite some spatial variability. Especially since one of the co-authors (I. Roghozina) is specialized in geothermal heat flow underneath ice sheets. Some evidence as to the choice of this value should be given as well as the eventual consequences of this choice.}

\todo{Develop this point.}

\revpoint{Line 199-200: 'Summer temperature and winter precipitation are most relevant to the glacier model as they drive summer melt and winter accumulation, respectively'.}

Moved 'respectively' to the end of the sentence.

\revpoint{Line 206: 're-projected to the Canadian ...'}

Added 'the'.

\revpoint{Line 216: spatially-smoothed}

Added a dash.

\revpoint{Line 276: However,}

Added a coma.

\revpoint{Line 284: to allow for comparison}

Added 'for'.

\revpoint{Line 294: compare modelled ice sheet ...}

Removed 'numerically.'

\revpoint{Line 362-367: This is a simple parameterization in Janssens and Huybrechts. If it would make such a difference, what is the rationale behind the fact that this has not been done? Please state.}

\todo{Develop this point.}

\revpoint{Figures: would it be possible to draw the limits of the Cordilleran ice sheets (as shown in Fig.~1) to the figures that show the simulated ice extents? It would make visual inspection easier.}

\todo{Develop this point.}

\end{itemize}

% ----------------------------------------------------------------------

\closing{Best regards,}

\end{letter}
\end{document}