### 2014-03-20 abstract.tex
Typesetting: title and authors

\documentclass[12pt]{article}
\usepackage[margin=25mm]{geometry}
\renewcommand{\familydefault}{\sfdefault}
\setlength\parindent{0pt}
\parskip=2ex
\thispagestyle{empty}
\begin{document}

\textbf{The effect of climate forcing on numerical simulations of the Cordilleran ice sheet at the Last Glacial Maximum}

J.~Seguinot\textsuperscript{1},
C.~Khroulev\textsuperscript{2},
I.~Rogozhina\textsuperscript{3},
A.~P.~Stroeven\textsuperscript{1},
and Q.~Zhang\textsuperscript{1}

\textsuperscript{1}Department of Physical Geography and Quaternary Geology, Stockholm University, Stockholm, Sweden \\
\textsuperscript{2}Geophysical Institute, University of Alaska Fairbanks, Fairbanks, AK, USA \\
\textsuperscript{3}Helmholtz Centre Potsdam, GFZ German Research Centre for Geosciences, Potsdam, Germany

\input{sections/abstract.tex}

\end{document}