### 2014-03-26 reviews/disc-ac2.tex

\documentclass[10pt]{article}

\usepackage[T1]{fontenc}
\usepackage[pdftex]{graphicx}
\usepackage[pdftex]{color}
\usepackage[colorlinks,citecolor=blue]{hyperref}
\usepackage{natbib}

\sloppy
\setlength{\parskip}{1.2ex}
\setlength{\parindent}{0em}
\renewcommand\familydefault{\sfdefault}

\graphicspath{{../figures/}}
\definecolor{journalname}{rgb}{0.34,0.59,0.82}

\begin{document}

\textbf{Authors' response to Referee {\#}2}
\bigskip

% ----------------------------------------------------------------------
% Interactive comment text begins
% ----------------------------------------------------------------------

\def\doi#1{doi:\allowbreak\href{http://dx.doi.org/#1}{#1}}
\def\referee#1{\bigskip\textcolor{journalname}{\textit{#1}}}

Dear Referee {\#}2,

Thank you very much for reading and positively commenting on our manuscript. We apologize for the lack of detail regarding isostatic issues, which we discuss below.

\referee{Before publication, I do have one scientific concern about the way the experiments were carried out. The model setup (Section 2) includes only brief mention of isostatic effects. These are really significant in a run of this sort because the elevation of the ice surface could be as much as $\sim$1800 m too high without the effect. Given that lapse rates are used to re-scale the climate data to elevations that are altered by the presence of the ice sheet, and that the sliding relation is strongly controlled by bedrock elevation, I think a more detailed discussion of isostacy is warranted. Specifically, given that the goal is an LGM configuration, I'd like assurance that the bedrock elevation has come close to equilibrium at the 10ka point that is reported on. This assurance should be simple to provide with minor modifications to the model setup section.}

The bedrock deformation model used in our study describes the flexure of an elastic lithosphere on top of an infinite half-space viscous mantle \citep{lingle-clark-1985}. It can be described by a single differential equation of the bed elevation $u(x, y, t)$,

\begin{equation}
    2 \nu |\nabla| \frac{\partial u}{\partial t}
    + \rho_r g u
    + D \nabla^4 u
    = \sigma_{zz},
\end{equation}

where $\nu=1\times10^{21}$\,Pa\,s is mantle viscosity, $\rho_r = 3300$\,kg\,m$^{-3}$ is lithosphere density, $D=5\times10^{24}$\,N\,m is the lithosphere's flexural rigidity, and $\sigma_{zz}$ corresponds to the ice load \citet{bueler-etal-2007}. In the left-hand part of this equation, the first term accounts for mantle relaxation, the second for point-wise isostasy, and the third for elastic flexure. Please refer to \citet{bueler-etal-2007} for a definition of the pseudo-differential operator $|\nabla|$. Due to high mantle viscosity $\nu$, there exist a time lag between the ice sheet growth and the isostatic bedrock response.

We would like to emphasize that most of our simulations do not reach a climatic equilibrium. In fact, simulations ran to climatic equilibrium (not shown in the paper) produce excessive ice cover in continental regions regardless of the forcing climatology used. Moreover, geomorphological and palaeoclimatic records give support for a short-lived Cordilleran ice sheet. Because the modelled ice sheet is in constant expansion, and because of the first term in the above equation, bedrock elevation does not come to equilibrium with the ice sheet configuration in our runs.

Although we do not believe that bedrock elevation must have been at equilibrium with the ice sheet configuration at the LGM, we recognize that through our simplistic palaeoclimatic approach, isostatic effects are another significant source of error on modelled ice surface elevation, which we omitted to discuss in our manuscript.

Through the till friction angle parameter $\phi$, sliding is controlled by bedrock elevation. This control is meant to reflect the presence of lubricating marine sediment at the ice sheet base. Such sediments are expected below current sea level, and, to a lesser extent, in areas that were submerged during former glacial periods, which is reflected in our simulation set-up by a linear increase of $\phi$ between 0 and 200 meters above sea level, where it reaches its maximum value.

We want to clarify here that $\phi$ is computed only once and for the entire simulation length at the beginning of the run, based on modern bedrock elevation. Because the distribution of marine sediments is not likely to change much during a single phase of ice sheet advance and retreat, sliding is not affected by bedrock elevation changes due to isostatic effects in our runs.

In general, it became clear to us through your comments that we need more details about computation of bedrock deformation and its effects in the manuscript. These will surely be added in the revised version.

\referee{The only writing problem I found was in figure 5, `logarythmic' should be `logarithmic'. Many thanks to the authors for getting the wording right the first time.}

Thank you for noticing this mistake. We will correct it.

\begin{thebibliography}{69}

\bibitem[{Lingle and Clark(1985)}]{lingle-clark-1985}
Lingle,~C.~S. and Clark,~J.~A.: A~numerical model of interactions between a~marine ice sheet and the solid Earth: application to a~West Antarctic ice stream, J. Geophys. Res., 90, 1100--1114, \doi{10.1029/JC090iC01p01100}, 1985.

\bibitem[{Bueler et~al.(2007)}]{bueler-etal-2007}
Bueler,~E. and Lingle,~C.~S. and Brown,~J.: Fast computation of a viscoelastic deformable Earth model for ice-sheet simulations, Ann. Glaciol., 46, 97--105, \doi{10.3189/172756407782871567}, 2007

\end{thebibliography}

% ----------------------------------------------------------------------
% Interactive comment text ends
% ----------------------------------------------------------------------

\end{document}