### 2014-06-27 figures/model-pdd.tex
Add PDD model algorythm from PISM docs.

\documentclass[border=5mm,tikz]{standalone}
\usepackage{amsmath}
\usetikzlibrary{chains}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}

\newcommand{\PDD}{\ensuremath{\text{PDD}}}
\newcommand{\Hs}{\ensuremath{H_{\text{snow}}}}
\newcommand{\Ms}{\ensuremath{\text{Melt}_{\text{snow}}}}
\newcommand{\Mi}{\ensuremath{\text{Melt}_{\text{ice}}}}

\begin{document}
  \begin{tikzpicture}[ auto, thick, font=\small, node distance=3mm,
    block/.style={rectangle, rounded corners, draw, fill=gray!5, text width=2in, text centered,
      minimum width=2in, minimum height=1cm},
    big block/.style={block, text width=3in},
    decision/.style={diamond, draw, fill=gray!5},
    line/.style={draw, -latex'},
    start chain=1 going below, every node/.style={on chain},
    every join/.style={-latex'},
    ]

    \node [big block] {Compute snow accumulation \mbox{``Accumulation''}, discarding rain};
    \node [big block, join] {Compute the number of \mbox{positive} degree days (\PDD) };
    \node [big block, join] {Update snow depth \mbox{$\Hs \gets \Hs + \text{Accumulation}$}};
    \node [big block, join] {Compute maximum possible snow melt \mbox{$\Ms^{\text{max}} := \PDD\times F_{s}$}};
    \node (melt all snow) [decision, join] {$\Ms^{\text{max}} > \Hs?$};

    \begin{scope}[start branch=yes, node distance=4em]
      \node (ice melt case) [block, on chain=going right] {\mbox{Melt all available snow:} $\Ms := \Hs$};
    \end{scope}
    \begin{scope}[start branch=no, node distance=4em]
      \node (no ice melt case) [block, on chain=going left] {Melt some of the snow: \mbox{$\Ms := \Ms^{\text{max}}$}};
    \end{scope}

    \node (excess pdds) [big block] {$\PDD_{\text{excess}} := \PDD - \Ms / F_{s}$};
    \node [big block, join] {Melt some ice: \mbox{$\Mi := \PDD_{\text{excess}} \times F_{i}$}};
    \node (refreeze ice) [decision, join] {Refreeze melted ice?};

    \begin{scope}[start branch=yes, node distance=4em]
      \node (refreeze both) [block, on chain=going right] {Refreeze is \mbox{$(\Ms + \Mi) \times \theta_{\text{refreeze}}$}};
    \end{scope}

    \begin{scope}[start branch=no, node distance=4em]
      \node (refreeze snow) [block, on chain=going left] {Refreeze is $\Ms \times \theta_{\text{refreeze}}$};
    \end{scope}

    \node (runoff) [big block] {$\text{Runoff} := \Ms + \Mi - \text{Refreeze}$};
    \node [big block, join] {Compute surface mass balance: $\text{SMB} := \text{Accumulation} - \text{Runoff}$};
    \node [big block, join] {Update snow depth: $\Hs \gets \Hs - \Ms$};
    % special edges:
    \path[line] (melt all snow) -- node {Yes} (ice melt case);
    \path[line] (melt all snow) -- node {No}  (no ice melt case);
    \path[line] (ice melt case) |- (excess pdds);
    \path[line] (no ice melt case) |- (excess pdds);
    \path[line] (refreeze ice) -- node {Yes} (refreeze both);
    \path[line] (refreeze ice) -- node {No}  (refreeze snow);
    \path[line] (refreeze both) |- (runoff);
    \path[line] (refreeze snow) |- (runoff);
  \end{tikzpicture}

\end{document}