### 2016-01-05 reviews/response-editor.tex
Initialize response to the editor.
This is just to save an email for later that we meant to send to the editor this morning if I had not overslept.

% response-marshall.tex
% ----------------------------------------------------------------------
\input{response-header}

\begin{document}
\textbf{Authors' response to the Editor}
\bigskip

Dear Hilmar Gudmundsson,
(cc co-authors and Copernicus publications)

As you probably noticed, Copernicus stopped peer review of our manuscript
"Numerical simulations of the Cordilleran ice sheet through the last glacial
cycle" when we were nearly ready to submit our responses. Since the deadline
coincided with new year celebrations, I could not obtain feedback from my
co-authors on time to finalize the responses.

We greatly apologize for this additional delay and ask for your assistance to
continue with the peer review that has required a large time investment from
our side and from the side of the reviewers. To some extent, the delay has been
caused by our wish to address the reviews as well as possible. In addition to a
change of my working environment, setting up and running the sensitivity study
asked by Alexander Jarosch on a different supercomputer, using a new allocation
that was meant for a different project and became effective only in October,
took us considerable time.

I contacted Svenja Lange prior to the submission deadline to inquire after the
possibility of a short-term extension but later realized that Copernicus
offices were closed for the new year celebration until yesterday. I have tried
to contact Natascha T\"{o}pfer yesterday (as suggested in the following email from
Copernicus) but then realised on the Cryosphere website that she is on a leave.
Is there a chance that the on-line submission system could be re-opened for us
to submit our responses?

Thank you very much in advance for your understanding,

Best regards,

Julien.

\end{document}