### 2016-11-04 cycle/abstract.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=25mm]{geometry}
\renewcommand{\familydefault}{\sfdefault}
\setlength\parindent{0pt}
\parskip=2ex
\thispagestyle{empty}
\begin{document}

% Title
\textbf{Numerical simulations of the Cordilleran ice sheet
        through the last glacial cycle}

% Authors
J.~Seguinot\textsuperscript{1,2,3},
I.~Rogozhina\textsuperscript{3},
A.~P.~Stroeven\textsuperscript{2},
M.~Margold\textsuperscript{2}
and J.~Kleman\textsuperscript{2}

% Affiliations
\textsuperscript{1}{Laboratory of Hydraulics, Hydrology and Glaciology,
                    ETH Z\"{u}rich, Z\"{u}rich, Switzerland}\\
\textsuperscript{2}{Department of Physical Geography and the
                    Bolin Centre for Climate Research,
                    Stockholm University, Stockholm, Sweden}\\
\textsuperscript{3}{Helmholtz Centre Potsdam,
                    GFZ German Research Centre for Geosciences,
                    Potsdam, Germany}\\

\begin{abstract}

After more than a century of geological research, the Cordilleran
ice sheet of North America remains among the least understood in
terms of its former extent, volume, and dynamics. Because of the
mountainous topography on which the ice sheet formed, geological
studies have often had only local or regional relevance, and shown
such a complexity that ice sheet-wide spatial reconstructions of
advance and retreat patterns are
lacking. Here we use a numerical ice sheet model calibrated against
field-based evidence to attempt a quantitative reconstruction of the
Cordilleran ice sheet history through the last glacial
cycle. A series of simulations is driven by time-dependent
temperature offsets from six proxy records located around the
globe. Although this approach reveals large variations in model
response to evolving climate forcing, all simulations produce two
major glaciations during marine oxygen isotope stages 4
(62.2--56.9\,ka) and 2 (23.2--16.9\,ka). The timing of
glaciation is better reproduced using temperature reconstructions
from Greenland and Antarctic ice cores than from regional oceanic
sediment cores. During most of the last glacial cycle, the modelled
ice cover is discontinuous and restricted to high mountain
areas. However, widespread precipitation over the Skeena Mountains
favours the persistence of a central ice dome throughout the glacial
cycle. It acts as a nucleation centre before the Last Glacial
Maximum and hosts the last remains of Cordilleran ice until the
middle Holocene (6.7\,ka).

\end{abstract}

\end{document}